---
title: "Australia GDP Forecasting"
output:
  html_document:
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

# Load packages
library(knitr)
library(zoo)
library(forecast)
library(vars)
library(lmtest)
library(sandwich)
library(tseries)
library(broom)
library(modelsummary)
library(readxl)
library(janitor)
library(scales)
library(glue)
library(fixest)
library(strucchange)
library(car)
library(multidplyr)
library(gridExtra)
library(marginaleffects)
library(tidyverse)

theme_set(theme_minimal(base_size = 12))

# Ensure directories
dir.create("output/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("output/tables", recursive = TRUE, showWarnings = FALSE)
dir.create("output/models", recursive = TRUE, showWarnings = FALSE)

# Paths
raw_path <- "data/data_raw.xlsx"
path_models <- "output/models/estimation.rds"
path_forecasts <- "output/models/forecasts.rds"

# Variables
sample_start <- "1993 Q1"
sample_end <- "2025 Q4"
evaluation_start <- "2016 Q1"
evaluation_end <- "2025 Q4"

extra_predictor <- c("unemp")
regressors <- c("gdp_yoy", "infl_yoy", "d_rate", extra_predictor)

mod_max_lags <- 4

hac_vcov <- function(fit, lag = 4) {
  NeweyWest(fit, lag = lag, prewhite = FALSE, adjust = TRUE)
}

# Flags
include_nowcasts <- FALSE

generate_models <- FALSE
generate_forecasts <- FALSE
```

# Build Dataset

```{r build_dataset}
# Raw data
clean_raw_data <- \(data) {
  res <- data %>%
    clean_names() %>%
    mutate(
      date = as.Date(date),
      quarter = as.yearqtr(date)
    )

  res
}

raw_data_quarterly <- read_xlsx(raw_path, sheet = "Q") %>%
  clean_raw_data() %>%
  select(quarter, rgdp, cpi)

raw_data_monthly <- read_xlsx(raw_path, sheet = "M") %>%
  clean_raw_data() %>%
  group_by(quarter) %>%
  summarise(unemp = mean(unemp), .groups = "drop")

raw_data_daily <- read_xlsx(raw_path, sheet = "D") %>%
  clean_raw_data() %>%
  group_by(quarter) %>%
  summarise(cash = mean(cash), .groups = "drop")

raw_inputs <- reduce(
  list(raw_data_quarterly, raw_data_monthly, raw_data_daily),
  ~ left_join(.x, .y, by = "quarter")
)

# Formatted data
dataset <- raw_inputs %>%
  arrange(quarter) %>%
  mutate(date = as.Date(quarter), .before = 1) %>%
  # Growth rates
  mutate(
    gdp_yoy = 100 * (log(rgdp) - log(lag(rgdp, 4))),
    infl_yoy = 100 * (log(cpi) - log(lag(cpi, 4))),
    rate = cash,
    d_rate = rate - lag(rate),
    d_unemp = unemp - lag(unemp),
  ) %>%
  # Filter sample
  filter(
    quarter >= sample_start,
    quarter <= sample_end
  ) %>%
  # drop_na(rgdp, cpi, cash, unemp, gdp_yoy, infl_yoy, rate) %>%
  mutate(t = 1:n()) %>%
  select(
    date,
    t,
    quarter,
    rgdp,
    cpi,
    cash,
    unemp,
    d_unemp,
    gdp_yoy,
    infl_yoy,
    rate,
    d_rate
  )

# Error if NAs in the data
invisible(na.fail(dataset))

kable(head(dataset, 8), digits = 3, caption = "Processed estimation dataset preview")

## Create lagged and filtered model dataset
model_dataset <- dataset %>%
  arrange(t) %>%
  reduce(seq(0, mod_max_lags), .init = ., \(data, lag) {
    res <- data %>%
      mutate(across(all_of(regressors), ~ lag(., lag), .names = "{.col}_l{lag}"))

    res
  }) %>%
    filter(t > mod_max_lags)

selection_dataset <- filter(model_dataset, quarter < evaluation_start)
```

# Descriptive Analysis

```{r summary_statistics}
analysis_variable_dictionary <- tribble(
    ~variable, ~label,
    "gdp_yoy", "GDP growth (YoY, %)",
    "infl_yoy", "Inflation (YoY, %)",
    "rate", "Cash rate (%)",
    "d_rate", "Change in cash rate (pp)",
    "unemp", "Unemployment rate (%)"
  )

# Summary statistics
summary_vars <- c("gdp_yoy", "infl_yoy", "rate", "d_rate", "unemp")

summary_stats <- dataset %>%
  select(all_of(summary_vars)) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
  group_by(variable) %>%
  summarise(
    n = n(),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    min = min(value, na.rm = TRUE),
    p25 = quantile(value, 0.25, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    p75 = quantile(value, 0.75, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(analysis_variable_dictionary, ., by = "variable")

kable(
  summary_stats,
  digits = 3,
  caption = "Summary statistics"
)
```

```{r series_plot, fig.width=10, fig.height=10}
# Series plot
plot_data <- dataset %>%
  select(date, all_of(summary_vars)) %>%
  pivot_longer(-date, names_to = "variable", values_to = "value") %>%
  left_join(analysis_variable_dictionary, by = "variable")

series_plot <- ggplot(plot_data, aes(x = date, y = value)) +
  geom_line(color = "#0b5d8c", linewidth = 0.6) +
  facet_wrap(~ label, scales = "free_y", ncol = 2) +
  labs(
    title = "Australia macroeconomic series",
    x = NULL, y = NULL
  ) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )

series_plot

# ggsave("output/figures/series_plot.png", series_plot, width = 10, height = 10, dpi = 300)
{series_plot + labs(title = NULL) } %>%
ggsave("output/figures/series_plot.png", ., width = 5, height = 3.5, dpi = 300)
```


```{r stationarity}
run_adf_urca <- \(x, variable, type = "drift", max_lags = 4) {
  x <- na.omit(x)

  test <- ur.df(
    x,
    type = type,          # "none", "drift", or "trend"
    lags = max_lags,
    selectlags = "AIC"
  )

  tau_name <- switch(
    type,
    "none" = "tau1",
    "drift" = "tau2",
    "trend" = "tau3"
  )

  stat <- test@teststat[1, tau_name]
  crit_5 <- test@cval[tau_name, "5pct"]
  crit_10 <- test@cval[tau_name, "10pct"]

  tibble(
    variable = variable,
    adf_type = type,
    adf_stat = stat,
    adf_crit_5pct = crit_5,
    adf_reject_unit_root_5pct = stat < crit_5,
    adf_crit_10pct = crit_10,
    adf_reject_unit_root_10pct = stat < crit_10
  )
}

adf_table <- map_dfr(summary_vars, ~ run_adf_urca(dataset[[.]], ., type = "drift"))

kable(
  adf_table,
  digits = 3,
  caption = "Augmented Dickey-Fuller results"
)
```

```{r acf_pacf_plot, fig.width=8, fig.height=8}
# ACF/PACF plot
get_acf_pacf_plot <- \(var, max_lag = 12, title = NULL) {
  acf_df <- var %>%
    acf(lag.max = max_lag, plot = FALSE) %>%
    tidy() %>%
    rename(ACF = acf)

  pacf_df <- var %>%
    pacf(lag.max = max_lag, plot = FALSE) %>%
    tidy() %>%
    rename(PACF = acf)

  acf_pacf_plot <- full_join(acf_df, pacf_df, by = "lag") %>%
    filter(lag > 0) %>%
    mutate(ci = qnorm((1 + 0.95)/2)/sqrt(length(var))) %>%
    pivot_longer(c(ACF, PACF), names_to = "type", values_to = "value") %>%
    ggplot(aes(x = lag, y = value)) +
    geom_hline(yintercept = 0, color = "grey50") +
    geom_segment(aes(xend = lag, yend = 0), color = "#c45508", linewidth = 0.7) +
    geom_hline(aes(yintercept = ci), linetype = "dashed") +
    geom_hline(aes(yintercept = -ci), linetype = "dashed") +
    facet_wrap(~ type, ncol = 1) +
    scale_x_continuous(breaks = pretty_breaks()) +
    labs(
      title = title,
      x = "Lag (quarters)",
      y = NULL
    )
}

acf_pacf_plot <- get_acf_pacf_plot(dataset$gdp_yoy, title = "GDP growth autocorrelation diagnostics")

acf_pacf_plot

# ggsave("output/figures/gdp_acf_pacf.png", acf_pacf_plot, width = 8, height = 8, dpi = 300)
{ acf_pacf_plot + labs(title = NULL)} %>%
ggsave("output/figures/gdp_acf_pacf.png", ., width = 6, height = 4.2, dpi = 300)
```

## Comment

### Variables

For the purposes of forecasting quarterly Y-o-Y GDP growth, the dataset is transformed
into quarterly variables. Real GDP and CPI are converted into year-on-year log
growth rates, while the cash rate is used mainly in first differences because
the level appears nonstationary. Unemployment is retained as the extra predictor.

For the training sample, we select 1993Q1–2015Q4.
For the evaluation sample, we use 2016Q1–2025Q4.

### Descriptive analysis

Australian GDP growth is fairly stable before COVID, but the 2020 collapse and 2021
rebound create extreme observations that dominate the later forecast evaluation.
Inflation and interest rates also shift sharply after 2021, so model stability and
out-of-sample performance are especially important.

### Stationarity and autocorrelation

The ADF tests support using GDP growth, inflation, unemployment and the change in
the cash rate as stationary variables. The ACF/PACF of GDP growth shows persistence,
especially at the first lag, which justifies dynamic specifications such as AR,
ARDL and VAR models.

## Breakpoint checking

```{r breakpoints}
get_test_models <- \(dataset, intercept = NULL, interaction = NULL) {
  if (!is.null(intercept)) {
    list(
      ar2 = glue("gdp_yoy ~ gdp_yoy_l1 + gdp_yoy_l2 + {intercept}"),
      ar4 = glue("gdp_yoy ~ gdp_yoy_l1 + gdp_yoy_l2 + gdp_yoy_l3 + gdp_yoy_l4 + {intercept}"),
      ardl2 = glue("gdp_yoy ~ {paste(c(unlist(map(regressors, ~ glue('{.}_l{seq(1, 2)}')))), collapse = '+')} + {intercept}"),
      ardl4 = glue("gdp_yoy ~ {paste(c(unlist(map(regressors, ~ glue('{.}_l{seq(1, 4)}')))), collapse = '+')} + {intercept}")
    ) %>%
      map(~ lm(formula(.), data = dataset))
  } else  if (!is.null(interaction)) {
    list(
      ar2 = glue("gdp_yoy ~ (gdp_yoy_l1 + gdp_yoy_l2) * {interaction}"),
      ar4 = glue("gdp_yoy ~ (gdp_yoy_l1 + gdp_yoy_l2 + gdp_yoy_l3 + gdp_yoy_l4) * {interaction}"),
      ardl2 = glue("gdp_yoy ~ ({paste(c(unlist(map(regressors, ~ glue('{.}_l{seq(1, 2)}')))), collapse = '+')}) * {interaction}"),
      ardl4 = glue("gdp_yoy ~ ({paste(c(unlist(map(regressors, ~ glue('{.}_l{seq(1, 4)}')))), collapse = '+')}) * {interaction}")
    ) %>%
      map(~ lm(formula(.), data = dataset))
  } else {
    list(
      ar2 = gdp_yoy ~ gdp_yoy_l1 + gdp_yoy_l2,
      ar4 = gdp_yoy ~ gdp_yoy_l1 + gdp_yoy_l2 + gdp_yoy_l3 + gdp_yoy_l4,
      ardl2 = reformulate(unlist(map(regressors, ~ glue("{.}_l{seq(1, 2)}"))), response = "gdp_yoy"),
      ardl4 = reformulate(unlist(map(regressors, ~ glue("{.}_l{seq(1, 4)}"))), response = "gdp_yoy")
    ) %>%
      map(~ lm(., data = dataset))
  }
}

# Pulse dummies
mod_resid <- list(
  ar2 = lm(gdp_yoy ~ gdp_yoy_l1 + gdp_yoy_l2, data = model_dataset),
  ar4 = lm(gdp_yoy ~ gdp_yoy_l1 + gdp_yoy_l2 + gdp_yoy_l3 + gdp_yoy_l4 , data = model_dataset),
  ardl2 = lm(reformulate(unlist(map(regressors, ~ glue("{.}_l{seq(1, 2)}"))), response = "gdp_yoy"), data = model_dataset),
  ardl4 = lm(reformulate(unlist(map(regressors, ~ glue("{.}_l{seq(1, 4)}"))), response = "gdp_yoy"), data = model_dataset)
)

## Check residuals (largest are 106, 110)
plot_residuals <- map_dfc(mod_resid, ~ rstandard(.)) %>%
  mutate(date = model_dataset$date) %>%
  pivot_longer(-date, names_to = "model", values_to = "res") %>%
  ggplot(aes(x = date, y = res)) +
  geom_line() +
  geom_hline(yintercept = c(-3, 3), color = "red") +
  geom_hline(yintercept = c(-2, 2), color = "red", linetype = "dashed") +
  facet_wrap(~ model) +
  labs(title = "Standardized residuals", x = NULL, y = NULL)

ggsave("output/figures/residuals.png", plot_residuals, width = 10, height = 10, dpi = 300)

# Breakpoints
breakpoint_dataset <- selection_dataset

mod_brkp <- get_test_models(breakpoint_dataset)

## Chow
shock_candidates <- tribble(
  ~name, ~start, ~comment,
  "gfc_2008q3", "2008 Q3", "GFC",
  "gfc_2008q4", "2008 Q4", "GFC",
  "gfc_2009q1", "2009 Q1", "GFC"
)

chow_test <- \(mod, t) {
  tryCatch(sctest(formula(mod), data = breakpoint_dataset, type = "Chow", point = t)$p.value, error = \(e) NA)
}

variance_break_test <- function(fit, break_idx) {
  e <- resid(fit)
  model_data <- model.frame(fit)

  idxs <- index(breakpoint_dataset$quarter)

  pre  <- e[idxs <=  break_idx]
  post <- e[idxs > break_idx]

  test <- var.test(post, pre)

  data.frame(
    var_pre = var(pre),
    var_post = var(post),
    F_stat = unname(test$statistic),
    p_value = test$p.value
  )
}

diag_chow <- shock_candidates %>%
  mutate(break_idx = match(start, breakpoint_dataset$quarter) - 1) %>%
  rowwise() %>%
  mutate(bind_cols(set_names(map(mod_brkp, ~ variance_break_test(., break_idx[[1]])$p_value), paste0("var_p_", names(mod_brkp))))) %>%
  mutate(bind_cols(set_names(map(mod_brkp, ~ chow_test(., break_idx[[1]])), paste0("chow_p_", names(mod_brkp)))))

kable(
  diag_chow,
  digits = 3,
  caption = "Chow tests for candidate breakpoints"
)

# GFC test
breakpoint_dataset <- breakpoint_dataset %>%
  mutate(
    post_gfc_2008q4 = as.integer(quarter >= "2008 Q4")
  )

mod_brkp_gfc <- get_test_models(breakpoint_dataset, intercept = "post_gfc_2008q4")

diagn_gfc <- map2_dfr(mod_brkp_gfc, names(mod_brkp_gfc), \(model, name) {
  # print(coeftest(model, vcov = hac_vcov(model)))

  test_int <- linearHypothesis(
    model,
    "post_gfc_2008q4",
    vcov. = hac_vcov(model),
    test = "Chisq"
  )

  tibble(model = name, int_p = test_int[2, 4])
})

mod_brkp_gfc_int <- get_test_models(breakpoint_dataset, interaction = "post_gfc_2008q4")

diagn_gfc_int <- map2_dfr(mod_brkp_gfc_int, names(mod_brkp_gfc_int), \(model, name) {
  # print(summary(model, vcov. = hac_vcov(model)))

  test_slop <- linearHypothesis(
    model,
    grep(":post_gfc_2008q4", names(coef(model)), value = TRUE),
    vcov. = hac_vcov(model),
    test = "Chisq"
  )

  tibble(
    model = name,
    slop_p = test_slop[2, 4],
  )
})

left_join(diagn_gfc, diagn_gfc_int, by = "model") %>%
  kable(
    digits = 3,
    caption = "Tests for GFC breakpoint intercept & slopes"
  )

## Recursive/CUSUM instability tests
diag_cusum <- map2_dfr(mod_brkp_gfc, names(mod_brkp_gfc), \(model, name) {
  cusum <- efp(formula(model), data = breakpoint_dataset, type = "Rec-CUSUM")
  test <- sctest(cusum)
  # print(test)

  tibble(
    model = name,
    cusum_p = test$p.value,
    series = list(
      tibble(
        process = c(rep(NA, cusum$nreg-1), cusum$process),
        boundary = c(rep(NA, cusum$nreg-1), strucchange::boundary(cusum, alpha = 0.05, alt.boundary = FALSE, functional = "max"))
      )
    )
  )
})

plot_cusum <- diag_cusum %>%
  unnest(series) %>%
  mutate(
    date = rep(breakpoint_dataset$date, length(mod_brkp_gfc)),
  ) %>%
  ggplot(aes(x = date, y = process)) +
  geom_line() +
  geom_line(aes(y = -boundary)) +
  geom_line(aes(y = boundary)) +
  geom_hline(aes(yintercept = 0)) +
  facet_wrap(~ model) +
  labs(title = "Recursive CUSUM test", x = NULL, y = NULL)

# ggsave("output/figures/cusum.png", plot_cusum, width = 10, height = 10, dpi = 300)
{ plot_cusum + labs(title = NULL) } %>%
ggsave("output/figures/cusum.png", ., width = 3.5, height = 3.5, dpi = 300)

# ## Bai-Perron (unknown dates)
walk2(mod_brkp_gfc, names(mod_brkp_gfc), \(model, name) {
  bp <- breakpoints(formula(model), data = breakpoint_dataset, h = 0.25)

  print(summary(bp))
  if (all(is.na(bp$breakpoints))) {
    print("No breakpoints detected")
  } else {
    print(breakpoint_dataset$quarter[bp$breakpoints])
  }
})
```

```{r breakpoint_plots, fig.width = 10, fig.height = 10}
plot_residuals
plot_cusum
```

### Comment

We use AR2, AR4, ARDL2, ARDL4 as general benchmarks for checking data/parameter stability.

We test for parameter instability using Chow tests at pre-specified GFC dates and
recursive CUSUM tests. The Chow tests indicate GFC-related instability around 2008Q3–2009Q1,
although variance tests suggest that this partly reflects a crisis-period volatility
shift rather than a clean coefficient break. Directly testing parameter instability
with a 2008Q4 GFC shift dummy indicates a statistically significant change in average
post-GFC output dynamics, while variable dynamics also changed for the ARDL models.
To preserve parsimonity in the main specifications, we include only the GFC shift dummy.
After its inclusion, the CUSUM paths remain within the 5% bands, providing no
strong evidence of remaining cumulative instability. Bai-Perron tests do not indicate
another unknown breakpoint.

The GFC period shows some evidence of instability, especially through changes in
residual variance and the intercept. A post-2008Q4 dummy is therefore included in
the main models to capture a persistent shift without overcomplicating the specification
with interaction terms.

```{r add_breakpoints}
model_dataset <- model_dataset %>%
  mutate(
    post_gfc_2008q4 = as.integer(quarter >= "2008 Q4"),
  )

selection_dataset <- filter(model_dataset, quarter < evaluation_start)
```

# Model estimation

## Estimation

### Comment

Specifications:

Benchmarks: AR2, VAR2 (gdp_yoy, infl_yoy, rate), VAR2 (gdp_yoy, infl_yoy, d_rate);

ARDL: iterating all variables & lags; easy to interpret

ARDL with seasonal ARMA errors (ARDL_ARMA): iterates over seasonal ARMA error specifications.

Later will see that basic ARDL errors have 4/8 quarter autocorrelation. This is
because we are modelling quarterly Y-o-Y growth, i.e., every quarter the current
GDP is compared to the GDP a year ago. This causes a "base effect", exemplified
by Covid: on 2020Q2 GDP dropped, resulting in a negative growth rate. Then, over
2020, GDP recovered, so in 2021Q2, the rebounded level of GDP is compared to the
2020Q2's low, causing a spike in the growth rate. This is also repeated for the
drop in 2021Q3 and spike in 2022Q3. Including lag-4 seasonal (AR)MA errors accounts
for this.

(S)ARMAX: same motivation as ARDL_ARMA, but the autoregressive component is specified
in the error term.


```{r estimate_models}
# Generate model zoo specifications
common_controls <- "post_gfc_2008q4"

model_specs_grid <- expand_grid(
  !!!setNames(
    rep(list(seq(-1, mod_max_lags)), length(regressors)),
    regressors
  ),
  ar = seq(0, mod_max_lags), ma = c(0, 1), s_ar = c(0, 1), s_ma = c(0,1)
) %>%
  filter(gdp_yoy != 0) %>%
  filter(!(ar != 0 & gdp_yoy != -1)) %>%
  filter(include_nowcasts | if_all(all_of(regressors), ~ . != 0)) %>%
  mutate(model = as.character(row_number()))

model_specs_ardl <- model_specs_grid %>%
  filter(if_all(c(ar, ma, s_ar, s_ma), ~ . == 0)) %>%
  rowwise() %>%
  mutate(regressors = list(reduce2(c_across(all_of(regressors)), regressors, \(acc, lag, var) {
    if (lag != -1) {
        min_lag <- 1 - as.numeric(include_nowcasts)
        c(acc, glue("{var}_l{seq(min_lag, lag)}"))
    } else acc
  }, .init = character(0)))) %>%
  mutate(family = "zoo", type = "ARDL") %>%
  mutate(controls = list(common_controls)) %>%
  mutate(model = paste0("ARDL", model))

model_specs_arma <- model_specs_grid %>%
  filter(if_any(c(ar, ma, s_ar, s_ma), ~ . != 0)) %>%
  rowwise() %>%
  mutate(regressors = list(reduce2(c_across(all_of(regressors)), regressors, \(acc, lag, var) {
    if (lag != -1) {
        min_lag <- 1 - as.numeric(include_nowcasts)
        c(acc, glue("{var}_l{seq(min_lag, lag)}"))
    } else acc
  }, .init = character(0)))) %>%
  mutate(family = "zoo", type = if_else(gdp_yoy == -1, "ARMAX", "ARDL_ARMA")) %>%
  mutate(controls = list(common_controls)) %>%
  mutate(model = paste0(if_else(gdp_yoy == -1, "ARMAX", "ARDL_ARMA"), model))

model_specs <- tribble(
  ~ model, ~ family, ~ type, ~ regressors, ~ controls, ~ gdp_yoy, ~ infl_yoy, ~ rate, ~ d_rate,
  "AR2", "benchmark", "ARDL",  c("gdp_yoy_l1", "gdp_yoy_l2"), character(0), 2, -1, -1, -1,
  "VAR2", "benchmark", "VAR", c("gdp_yoy", "infl_yoy", "rate"), character(0), 2, 2, 2, -1,
  "VAR2d", "benchmark", "VAR", c("gdp_yoy", "infl_yoy", "d_rate"), character(0), 2, 2, -1, 2
) %>%
  bind_rows(model_specs_ardl) %>%
  bind_rows(model_specs_arma) %>%
  mutate(
    across(-c(model, family, type, regressors, controls, ar, ma, s_ar, s_ma), ~ replace_na(., -1)),
    across(c(ar, ma, s_ar, s_ma), ~ replace_na(., 0))
  )

# Model estimation
## ARDL
fit_lm_model <- function(spec, data) {
  all_regressors <- c(spec$regressors[[1]], spec$controls[[1]])

  model_formula <- reformulate(all_regressors, response = "gdp_yoy")

  fit <- lm(model_formula, data = data)

  tibble(
    fit = list(fit)
  )
}

## VAR
fit_var_model <- function(spec, data, lag_order = 2L) {
  fit_data <- bind_rows(
    dataset[seq(mod_max_lags - lag_order + 1, mod_max_lags),],
    data
  )

  endogenous <- spec$regressors[[1]]
  controls <- spec$controls[[1]]

  if (length(controls) > 0) {
    fit <- VAR(
      y = fit_data[, endogenous],
      p = lag_order,
      type = "const",
      exogen = fit_data[, controls]
    )
  } else {
    fit <- VAR(
      y = fit_data[, endogenous],
      p = lag_order,
      type = "const"
    )
  }

  tibble(fit = list(fit))
}

## ARMA
fit_arma_model <- function(spec, data) {
  all_regressors <- c(spec$regressors[[1]], spec$controls[[1]])

  max_lag <- max(spec$ar, spec$ma, spec$s_ma, spec$s_ar)

  y <- data$gdp_yoy

  fit <- tryCatch({
    if (length(all_regressors) > 0) {
      Arima(
        y, xreg = as.matrix(data[, all_regressors]),
        order = c(spec$ar, 0, spec$ma),
        seasonal = list(order = c(spec$s_ar, 0, spec$s_ma), period = 4),
        method = "ML"
      )
    } else {
      Arima(
        y,
        order = c(spec$ar, 0, spec$ma),
        seasonal = list(order = c(spec$s_ar, 0, spec$s_ma), period = 4),
        method = "ML"
      )
    }
  }, error = \(e) {
    message(paste("Error for model", spec$model[[1]]))
    stop(e)
  })

  tibble(fit = list(fit))
}

## Iterate over all specs
if (generate_models | !file.exists(path_models)) {
  cl <- new_cluster(parallel::detectCores()-1)
  cluster_copy(cl, c(
    "fit_lm_model", "fit_var_model", "fit_arma_model",
    "selection_dataset", "dataset", "mod_max_lags"
  ))
  cluster_library(cl, c("dplyr", "forecast", "vars"))

  model_results <- model_specs %>%
    rowwise() %>%
    partition(cl) %>%
    mutate(switch(type,
      "ARDL" = fit_lm_model(cur_data(), selection_dataset),
      "VAR" = fit_var_model(cur_data(), selection_dataset),
      "ARMAX" = fit_arma_model(cur_data(), selection_dataset),
      "ARDL_ARMA" = fit_arma_model(cur_data(), selection_dataset)
    )) %>%
    collect() %>%
    ungroup()

  rm(cl)

  saveRDS(model_results, path_models)
} else {
  model_results <- readRDS(path_models)
}
```

## Shortlisting

```{r shortlisting}
# Model selection

## ICs
model_diagn_ic <- model_results %>%
  rowwise() %>%
  mutate(across(fit, list(bic = BIC, aic = AIC))) %>%
  ungroup()

model_diagn_ic %>%
  filter(family == "zoo") %>%
  mutate(group = if_else(type %in% c("ARDL", "ARDL_ARMA"), gdp_yoy, ar)) %>%
  ggplot(aes(x = fit_bic, y = fit_aic, color = group)) + geom_point() + facet_wrap(~ type)

### Can see that models without gdp_yoy are dominated

model_shortlist <- model_diagn_ic %>%
  filter(family == "benchmark") %>%
  mutate(selection = "benchmark") %>%
  bind_rows(
    bind_rows(
      model_diagn_ic %>%
        slice_min(fit_bic, n = 5, with_ties = FALSE, by = type) %>%
        mutate(selection = "Best BIC"),
      model_diagn_ic %>%
        slice_min(fit_aic, n = 5, with_ties = FALSE, by = type) %>%
        mutate(selection = "Best AIC")
    ) %>% arrange(type, selection)
  ) %>%
  distinct(pick(-selection), .keep_all = TRUE)

model_shortlist %>%
  select(-c(type, regressors, controls, fit)) %>%
  kable()
```

### Comment

We shortlist models with BIC & AIC, top 5 for each type for IC, so that model choice is
not driven by one criterion.
Models without an autoregressive component are dominated, and both BIC&AIC prefer
models with at least seasonal MA errors.
BIC tends to select simpler models with fewer variables and lags, while AIC favours
larger models with more lags or ARMA errors.
Note: VAR ICs cannot be compared with non-VAR models as they are for the whole model,
not just the GDP part.

## Diagnostics

```{r diagnostics}
# Diagnostics

## ARDL
ar_root_check <- function(fit, prefix = "gdp_yoy_l") {
  b <- coef(fit)

  phi <- b[str_detect(names(b), paste0("^", prefix))]
  phi <- phi[order(as.integer(str_extract(names(phi), "\\d+$")))]

  if (length(phi) == 0) return(NA)

  roots <- polyroot(c(1, -phi))

  tibble(
    min_abs_root = min(Mod(roots)),
    stable = min(Mod(roots)) > 1
  )
}

extract_lm_diagnostics <- function(model_info, data) {
  fit <- model_info$fit[[1]]

  e <- resid(fit)

  diagn_sign <- map_dfc(regressors, \(reg) {
    if (str_detect(paste(names(coef(fit)), collapse = " "), reg)) {
      tibble(
        "sum_coef_{reg}" := sum(coef(fit)[grep(reg, names(coef(fit)), value = TRUE)]),
        "joint_{reg}_p" := linearHypothesis(fit, grep(reg, names(coef(fit)), value = TRUE), vcov. = hac_vcov(fit), test = "Chisq")[[4]][[2]]
      )
    } else tibble("sum_coef_{reg}" := NA, "joint_{reg}_p" := NA)
  })

  cusum <- efp(fit, data = data, type = "Rec-CUSUM")

  # Plots
  ## Residuals
  resid_plot <- tibble(
    res = rstandard(fit),
    date = data$date
  ) %>%
    ggplot(aes(x = date, y = res)) +
    geom_line() +
    geom_hline(yintercept = c(-3, 3), color = "red") +
    geom_hline(yintercept = c(-2, 2), color = "red", linetype = "dashed") +
    labs(title = glue("Standardized residuals for {model_info$model}"), x = NULL, y = NULL)

  ## Residual autocorrelation
  acf_pacf_plot <- get_acf_pacf_plot(e, title = glue("Residual ACF/PACF for {model_info$model}"))

  ## Recursive CUSUM
  cusum_plot <- tibble(
    process = c(rep(NA, cusum$nreg-1), cusum$process),
    boundary = c(rep(NA, cusum$nreg-1), strucchange::boundary(cusum, alpha = 0.05, alt.boundary = FALSE, functional = "max")),
    date = data$date,
  ) %>%
    ggplot(aes(x = date, y = process)) +
    geom_line() +
    geom_line(aes(y = -boundary)) +
    geom_line(aes(y = boundary)) +
    geom_hline(aes(yintercept = 0)) +
    labs(title = glue("Recursive CUSUM test for {model_info$model}"), x = NULL, y = NULL)

  # Test values
  diagn_main <- tibble(
    # Residual autocorrelation
    diagn_bg_4_p = bgtest(fit, order = 4)$p.value,
    diagn_bg_8_p = bgtest(fit, order = 8)$p.value,
    diagn_lb_4_p = Box.test(e, lag = 4, type = "Ljung-Box", fitdf = model_info$gdp_yoy)$p.value,
    diagn_lb_8_p = Box.test(e, lag = 8, type = "Ljung-Box", fitdf = model_info$gdp_yoy)$p.value,
    # Heteroskedasticity (Breusch-Pagan)
    diagn_het_p = bptest(fit)$p.value,
    # Normality
    diagn_jb_p = jarque.bera.test(residuals(fit))$p.value,
    # Stationarity (for AR part) (should be > 1)
    diagn_ar_root = ar_root_check(fit)$min_abs_root,
    # Parameter stability
    diagn_cusum_p = sctest(cusum)$p.value,
    # Block t-tests
    !!!diagn_sign,
    # Plots
    resid_plot = list(resid_plot),
    acf_pacf_plot = list(acf_pacf_plot),
    cusum_plot = list(cusum_plot)
  )

  diagn_main
}

## VAR
extract_var_diagnostics <- function(model_info, data) {
  fit <- model_info$fit[[1]]

  oth_vars <- names(fit$varresult)[-1]

  # Refit with (Intercept) instead of const, so NeweyWest works
  gdp_eq <- update(fit$varresult$gdp_yoy, . ~ . - const + 1, data = fit$varresult$gdp_yoy$model)

  cusum <- stability(fit, type = "Rec-CUSUM")$stability$gdp_yoy

  ## Recursive CUSUM
  cusum_plot <- tibble(
    process = c(rep(NA, cusum$nreg-1), cusum$process),
    boundary = c(rep(NA, cusum$nreg-1), strucchange::boundary(cusum, alpha = 0.05, alt.boundary = FALSE, functional = "max")),
    date = data$date,
  ) %>%
    ggplot(aes(x = date, y = process)) +
    geom_line() +
    geom_line(aes(y = -boundary)) +
    geom_line(aes(y = boundary)) +
    geom_hline(aes(yintercept = 0)) +
    labs(title = glue("Recursive CUSUM test for {model_info$model}"), x = NULL, y = NULL)

  diagn_sign <- map_dfc(model_info$regressors[[1]], \(reg) {
    if (str_detect(paste(names(coef(gdp_eq)), collapse = " "), reg)) {
      tibble(
        "sum_coef_{reg}" := sum(coef(gdp_eq)[grep(reg, names(coef(gdp_eq)), value = TRUE)]),
        "joint_{reg}_p" := linearHypothesis(gdp_eq, grep(reg, names(coef(gdp_eq)), value = TRUE), vcov. = hac_vcov(gdp_eq), test = "Chisq")[[4]][[2]]
      )
    } else tibble("sum_coef_{reg}" := NA, "joint_{reg}_p" := NA)
  })

  # diagn_granger <- c(
  #   map_dbl(oth_vars, ~ causality(fit, cause = .)$Granger$p.value[[1]]),
  #   causality(fit, cause = oth_vars)$Granger$p.value[[1]]
  # ) %>%
  #   setNames(c(paste0("gr_", oth_vars, "_p"), "gr_joint_p"))

  diagn_main <- tibble(
    # Residual correlation
    diagn_bg_4_p = serial.test(fit, lags.pt = 4, type = "BG")$serial$p.value,
    diagn_bg_8_p = serial.test(fit, lags.pt = 8, type = "BG")$serial$p.value,
    # Heteroskedasticity (ARCH)
    diagn_het_p = arch.test(fit, lags.multi = 4)$arch.mul$p.value,
    # Normality
    diagn_jb_p = normality.test(fit)$jb.mul$JB$p.value[[1]],
    # Stationarity
    diagn_var_eigen = max(roots(fit)),
    # Recursive CUSUM,
    diagn_cusum_p = sctest(cusum)$p.value,
    cusum_plot = list(cusum_plot),
    # # Granger causality
    # !!!diagn_granger
    # Effects and joint significance in GDP equation
    !!!diagn_sign
  )

  diagn_main
}

## ARMA
arma_root_check <- function(fit, prefix = "gdp_yoy_l") {
  b <- coef(fit)
  names_b <- names(b)

  # Non-seasonal dependent lag coefficients: ar1, ar2, ...
  ar_lag <- b[str_detect(names_b, glue("^{prefix}[0-9]+$"))]
  ar_lag <- ar_lag[order(as.integer(str_extract(names(ar_lag), "\\d+$")))]

  # Non-seasonal AR coefficients: ar1, ar2, ...
  ar <- b[str_detect(names_b, "^ar[0-9]+$")]
  ar <- ar[order(as.integer(str_extract(names(ar), "\\d+$")))]

  # Non-seasonal MA coefficients: ma1, ma2, ...
  ma <- b[str_detect(names_b, "^ma[0-9]+$")]
  ma <- ma[order(as.integer(str_extract(names(ma), "\\d+$")))]

  # Seasonal AR coefficients: sar1, sar2, ...
  sar <- b[str_detect(names_b, "^sar[0-9]+$")]
  sar <- sar[order(as.integer(str_extract(names(sar), "\\d+$")))]

  # Seasonal MA coefficients: sma1, sma2, ...
  sma <- b[str_detect(names_b, "^sma[0-9]+$")]
  sma <- sma[order(as.integer(str_extract(names(sma), "\\d+$")))]

  ord <- arimaorder(fit)
  period <- ord[["Frequency"]]

  make_seasonal_poly <- function(coefs, sign = -1, period = 4) {
    if (length(coefs) == 0) return(NULL)

    poly <- rep(0, length(coefs) * period + 1)
    poly[1] <- 1

    for (i in seq_along(coefs)) {
      poly[1 + i * period] <- sign * coefs[i]
    }

    poly
  }

  ar_lag_poly <- if (length(ar_lag) > 0) c(1, -ar_lag) else NULL
  ar_poly <- if (length(ar) > 0) c(1, -ar) else NULL
  ma_poly <- if (length(ma) > 0) c(1,  ma) else NULL

  sar_poly <- make_seasonal_poly(sar, sign = -1, period = period)
  sma_poly <- make_seasonal_poly(sma, sign =  1, period = period)

  ar_roots <- c(
    if (!is.null(ar_lag_poly)) polyroot(ar_lag_poly) else numeric(0),
    if (!is.null(ar_poly)) polyroot(ar_poly) else numeric(0),
    if (!is.null(sar_poly)) polyroot(sar_poly) else numeric(0)
  )

  ma_roots <- c(
    if (!is.null(ma_poly)) polyroot(ma_poly) else numeric(0),
    if (!is.null(sma_poly)) polyroot(sma_poly) else numeric(0)
  )

  tibble(
    ar_min_root = if (length(ar_roots) > 0) min(Mod(ar_roots)) else NA,
    ma_min_root = if (length(ma_roots) > 0) min(Mod(ma_roots)) else NA,
  )
}

wald_arima_block <- function(fit, pattern) {
  b <- coef(fit)
  V <- fit$var.coef

  idx <- str_detect(names(b), pattern)

  if (!any(idx)) {
    return(p_value = NA)
  }

  b_block <- b[idx]
  V_block <- V[idx, idx, drop = FALSE]

  # robust to near-singular covariance matrices
  stat <- tryCatch(
    as.numeric(t(b_block) %*% solve(V_block, b_block)),
    error = function(e) as.numeric(t(b_block) %*% MASS::ginv(V_block) %*% b_block)
  )

  df <- length(b_block)

  tibble(
    eff = sum(b_block),
    p_value = pchisq(stat, df = df, lower.tail = FALSE)
  )
}

extract_arma_diagnostics <- \(model_info, data) {
  fit <- model_info$fit[[1]]
  model_name <- model_info$model

  e <- as.numeric(residuals(fit))
  e <- e[is.finite(e)]
  sigma_hat <- sqrt(fit$sigma2)

  # ARIMA order
  ord <- arimaorder(fit)

  arma_df <- sum(ord[c("p", "q", "P", "Q")], na.rm = TRUE)

  safe_lb <- function(e, lag, fitdf = 0) {
    if (lag <= fitdf) return(NA)
    Box.test(e, lag = lag, type = "Ljung-Box", fitdf = fitdf)$p.value
  }

  # ARCH test fallback: Ljung-Box on squared residuals
  arch_lb <- function(e, lag = 4) {
    Box.test(e^2, lag = lag, type = "Ljung-Box")$p.value
  }

  # Residual plots
  resid_plot <- tibble(
    res = as.numeric(e) / sigma_hat,
    date = tail(data$date, length(e))
  ) %>%
    ggplot(aes(x = date, y = res)) +
    geom_line() +
    geom_hline(yintercept = c(-3, 3), color = "red") +
    geom_hline(yintercept = c(-2, 2), color = "red", linetype = "dashed") +
    labs(
      title = glue("Standardized residuals for {model_name}"),
      x = NULL,
      y = NULL
    )

  acf_pacf_plot <- get_acf_pacf_plot(
    e,
    title = glue("Residual ACF/PACF for {model_name}")
  )

  root_info <- arma_root_check(fit)

  diagn_sign <- map_dfc(c(regressors, "^ar"), \(reg) {
     if (str_detect(paste(names(coef(fit)), collapse = " "), reg)) {
      test <- wald_arima_block(fit, reg)
      tibble(
        "sum_coef_{reg}" := test$eff,
        "joint_{reg}_p" := test$p_value
      )
     } else tibble("sum_coef_{reg}" := NA, "joint_{reg}_p" := NA)
  }) %>%
    mutate(
      sum_coef_gdp_yoy = coalesce(sum_coef_gdp_yoy, `sum_coef_^ar`),
      joint_gdp_yoy_p = coalesce(joint_gdp_yoy_p, `joint_^ar_p`),
      .keep = "unused"
    )

  tibble(
    diagn_lb_4_p = safe_lb(e, lag = 4, fitdf = arma_df),
    diagn_lb_8_p = safe_lb(e, lag = 8, fitdf = arma_df),

    diagn_het_p = arch_lb(e, lag = 4),

    diagn_jb_p = jarque.bera.test(e)$p.value,

    diagn_ar_root = root_info$ar_min_root,
    diagn_ma_root = root_info$ma_min_root,

    !!!diagn_sign,

    resid_plot = list(resid_plot),
    acf_pacf_plot = list(acf_pacf_plot)
  )
}

model_diagnostics <- model_shortlist %>%
  rowwise() %>%
  mutate(switch(
    type,
    "ARDL" = extract_lm_diagnostics(cur_data(), selection_dataset),
    "VAR" = extract_var_diagnostics(cur_data(), selection_dataset),
    "ARDL_ARMA" = extract_arma_diagnostics(cur_data(), selection_dataset),
    "ARMAX" = extract_arma_diagnostics(cur_data(), selection_dataset)
  )) %>%
  ungroup()

model_diagnostics %>%
  select(-c(type, regressors, controls, fit, fit_aic, fit_bic, selection, ends_with("plot"), starts_with("sum_coef_"), starts_with("joint_"), contains("gr"))) %>%
  mutate(diagn_ma_root = sprintf("%.8f", diagn_ma_root)) %>%
  kable(
    digits = 3,
    caption = "Model diagnostics"
  )

model_diagnostics %>%
  select(-c(type, regressors, controls, fit, fit_aic, fit_bic, selection, ends_with("plot"), starts_with("diagn_"))) %>%
  kable(
    digits = 3,
    caption = "Model economic hypotheses"
  )
```

```{r diagnostics_plots, fig.width = 25, fig.height = 25, out.width="100%"}
diagn_resid_plot <- grid.arrange(grobs = discard(model_diagnostics$resid_plot, is.null))
ggsave("output/figures/diagn_std_residuals.png", diagn_resid_plot, width = 25, height = 20, dpi = 300)

diagn_acf_pacf_plot <- grid.arrange(grobs = discard(model_diagnostics$acf_pacf_plot, is.null))
ggsave("output/figures/diagn_acf_pacf.png", diagn_acf_pacf_plot, width = 25, height = 20, dpi = 300)

diagn_cusum_plot <- grid.arrange(grobs = discard(model_diagnostics$cusum_plot, is.null))
ggsave("output/figures/diagn_cusum.png", diagn_cusum_plot, width = 25, height = 20, dpi = 300)
```

### Comment

#### Residual diagnostics

The plain ARDL models improve on the basic benchmarks (heteroskedasticity, normality,
stability) but still show residual autocorrelation (base effects) and some instability,
so they are not fully satisfactory as final models.

The ARDL_ARMA and ARMAX models perform better on residual autocorrelation tests,
although they have almost unit MA roots. This means that they should be interpreted
mainly as forecasting models rather than structural economic models.

#### Economic hypothesis

`sum_coef` are lagged variable coefficient sums; interpreted mostly for their sign.
`joint_p` is the p-value for the null hypothesis that all lags are 0.

The block tests for variable lags suggest that GDP persistence (positive), inflation
(negative) and unemployment (positive) contain useful predictive information for
GDP growth.
Remember that the effects are not causal; the positive unemployment effect likely
reflects cyclical mean reversion, where higher unemployment occurs during recessions
followed by recoveries.
The evidence for the relationship with the (change in) the cash rate is mixed both
in terms of sign and significance, depending on the specification.


# Recursive Forecasts

## Forecasting

```{r recursive_forecasts}
# Run forecasts
generate_recursive_forecast <- function(spec, data, eval_start, eval_end) {
  eval_quarters <- data %>%
    filter(
      quarter >= eval_start,
      quarter <= eval_end
    ) %>%
    pull(quarter)

  map_dfr(
    eval_quarters,
    function(target_quarter) {
      # browser()
      train_data <- data %>% filter(quarter < target_quarter)
      target_row <- data %>% filter(quarter == target_quarter)

      if (nrow(train_data) == 0 || nrow(target_row) != 1) {
        return(tibble())
      }

      model_info <- switch(spec$type,
        "ARDL" = fit_lm_model(spec, train_data),
        "VAR" = fit_var_model(spec, train_data),
        "ARDL_ARMA" = fit_arma_model(spec, train_data),
        "ARMAX" = fit_arma_model(spec, train_data)
      )

      forecast_value <- if (spec$type == "ARDL") {
        predict(model_info$fit[[1]], newdata = target_row)
      } else if (spec$type == "VAR") {
        predict(model_info$fit[[1]], n.ahead = 1)$fcst$gdp_yoy[, "fcst"]
      } else if (spec$type %in% c("ARDL_ARMA", "ARMAX")) {
        all_regressors <- c(spec$regressors[[1]], spec$controls[[1]])
        as.double(forecast(model_info$fit[[1]], xreg = as.matrix(target_row[, all_regressors]), h = 1)$mean)
      }

      actual_value <- target_row$gdp_yoy[[1]]
      error_value <- actual_value - forecast_value

      tibble(
        date = target_row$date[[1]],
        quarter = target_quarter,
        fit = list(model_info$fit[[1]]),
        forecast = forecast_value,
        actual = actual_value,
        error = error_value,
      )
    }
  )
}

if (generate_forecasts | !file.exists(path_forecasts)) {
  forecast_results <- model_shortlist %>%
    rowwise() %>%
    mutate(frcst = list(generate_recursive_forecast(cur_data(), model_dataset, evaluation_start, evaluation_end)))

  saveRDS(forecast_results, path_forecasts)
} else {
  forecast_results <- readRDS(path_forecasts)
}
```

## Forecast evaluation

```{r forecast_eval}
# Evaluation metrics
get_act_vs_frcst_plot <- function(forecast_results, title = "Recursive 1-ahead forecasts") {
  forecast_results %>%
  ggplot(aes(x = date)) +
    geom_col(aes(y = error, fill = "Error")) +
    geom_line(aes(y = actual, color = "Actual")) +
    geom_line(aes(y = forecast, color = "Forecast")) +
    scale_color_manual(values = c("Actual" = "black", "Forecast" = "red", "Error" = "grey")) +
    scale_fill_manual(values = c("Error" = "grey")) +
    labs(
      title = title,
      x = NULL,
      y = "GDP growth (YoY, %)",
      color = NULL, fill = NULL
    ) +
    theme(panel.grid.minor = element_blank())
}

get_cumse_plot <- function(forecast_results, title = "Cumulative squared forecast errors") {
  forecast_results %>%
    mutate(cumulative_squared_error = cumsum(error^2)) %>%
    ggplot(aes(x = date, y = cumulative_squared_error)) +
    geom_line() +
    labs(
      title = title,
      x = NULL,
      y = "Cumulative squared error"
    ) +
    theme(panel.grid.minor = element_blank())
}


get_model_forecast_eval <- \(model_info, spec_ar2, spec_var2) {
  fit_data <- model_info$frcst[[1]]

  fit_err <- lm(error ~ 1, data = fit_data)

  if (model_info$model != "AR2") {
    dm_ar2_p <- dm.test(fit_data$error, spec_ar2$frcst[[1]]$error, alternative = "less")$p.value

    diff_loss_ar2 <- fit_data$error^2 - spec_ar2$frcst[[1]]$error^2
    fit_loss_ar2 <- lm(diff_loss_ar2 ~ 1)

    # hac_loss_ar2_p <- coeftest(fit_loss_ar2, vcov. = hac_vcov(fit_loss_ar2, lag = 4))[1, "Pr(>|t|)"]
    hac_loss_ar2_p <- hypotheses(fit_loss_ar2, "`(Intercept)` >= 0", vcov = hac_vcov(fit_loss_ar2, lag = 4))$p.value
  } else {
    dm_ar2_p <- NA
    hac_loss_ar2_p <- NA
  }


  if (model_info$model != "VAR2") {
    dm_var2_p <- dm.test(fit_data$error, spec_var2$frcst[[1]]$error, alternative = "less")$p.value

    diff_loss_var2 <- fit_data$error^2 - spec_var2$frcst[[1]]$error^2
    fit_loss_var2 <- lm(diff_loss_var2 ~ 1)

    # hac_loss_var2_p <- coeftest(fit_loss_var2, vcov. = hac_vcov(fit_loss_var2, lag = 4))[1, "Pr(>|t|)"]
    hac_loss_var2_p <- hypotheses(fit_loss_var2, "`(Intercept)` >= 0", vcov = hac_vcov(fit_loss_var2, lag = 4))$p.value
  } else {
    dm_var2_p <- NA
    hac_loss_var2_p <- NA
  }

  tibble(
    # Bias test: H0 E(error) = 0
    bias_p = coeftest(fit_err, vcov. = hac_vcov(fit_err, lag = 4))[1, "Pr(>|t|)"],
    # Weak efficiency
    error_bg4_p = bgtest(fit_err, order = 4)$p.value,
    error_lb4_p = Box.test(fit_data$error, lag = 4, type = "Ljung-Box")$p.value,
    error_lb8_p = Box.test(fit_data$error, lag = 8, type = "Ljung-Box")$p.value,
    # Diebold-Mariano
    dm_ar2_p = dm_ar2_p,
    dm_var2_p = dm_var2_p,
    # HAC loss tests
    hac_loss_ar2_p,
    hac_loss_var2_p
  )
}

## Test full sample and also excluding covid (2020Q1-2022Q4)
forecast_metrics <- forecast_results %>%
  rowwise() %>%
  mutate(
    summarise(
      frcst,
      mfe = mean(error, na.rm = TRUE),
      msfe = mean(error^2, na.rm = TRUE),
      rmsfe = sqrt(msfe),
      mafe = mean(abs(error), na.rm = TRUE),
      act_frcst_plot = list(get_act_vs_frcst_plot(frcst, title = glue("Rec. 1-ahead frcst for {model}"))),
      cumse_plot = list(get_cumse_plot(frcst, title = glue("Cum. sq. frcst errors for {model}")))
    )
  ) %>%
  mutate(
    summarise(
      filter(frcst, date < "2020-01-01" | date > "2022-10-01"),
      msfe_exc = mean(error^2, na.rm = TRUE),
      rmsfe_exc = sqrt(msfe_exc),
      mafe_exc = mean(abs(error), na.rm = TRUE),
      cumse_exc_plot = list(get_cumse_plot(cur_data(), title = glue("Cum. sq. frcst errors for {model} (excl. COVID)")))
    )
  ) %>%
  ungroup() %>%
  mutate(across(c(msfe, msfe_exc), ~ . / .[model == "AR2"], .names = "{.col}_v_AR2")) %>%
  mutate(across(c(msfe, msfe_exc), ~ . / .[model == "VAR2"], .names = "{.col}_v_VAR2")) %>%
  rowwise() %>%
  mutate(get_model_forecast_eval(
    cur_data(),
    spec_ar2 = .[.$model == "AR2",],
    spec_var2 = .[.$model == "VAR2",]
  ))

forecast_metrics %>%
  select(-c(type, regressors, controls, fit, fit_bic, fit_aic, selection, frcst, ends_with("plot"))) %>%
  kable(
    digits = 3,
    caption = "Forecast performance"
  )
```

```{r forecast_eval_plots, fig.width = 30, fig.height = 20, out.width="100%"}
eval_act_frcst_plot_comb <- forecast_metrics %>%
  mutate(class = if_else(family == "benchmark", glue("Benchmark ({model})"), type)) %>%
  select(model, type, class, frcst) %>%
  unnest(frcst) %>%
  ggplot(aes(x = date)) +
    geom_line(aes(y = actual, color = "Actual", linewidth = "Actual", linetype = "Actual")) +
    geom_line(aes(y = forecast, group = model, color = class, linewidth = class, linetype = class)) +
    scale_color_manual(
      values = c(
        "Actual" = "#111111",
        "Benchmark (AR2)" = "#4c78a8",
        "Benchmark (VAR2)" = "#f58518",
        "Benchmark (VAR2d)" = "#f58518",
        "ARDL" = "#72b7b2",
        "ARDL_ARMA" = "#e45756",
        "ARMAX" = "violet"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "Actual" = 1.5,
        "Benchmark (AR2)" = 1.5,
        "Benchmark (VAR2)" = 1.5,
        "Benchmark (VAR2d)" = 1.5,
        "ARDL" = 0.7,
        "ARDL_ARMA" = 0.7,
        "ARMAX" = 0.7
      )
    ) +
    scale_linetype_manual(
      values = c(
        "Actual" = "solid",
        "Benchmark (AR2)" = "solid",
        "Benchmark (VAR2)" = "solid",
        "Benchmark (VAR2d)" = "dashed",
        "ARDL" = "solid",
        "ARDL_ARMA" = "solid",
        "ARMAX" = "solid"
      )
    ) +
    labs(
      title = "Recursive 1-ahead forecasts",
      x = NULL,
      y = "GDP growth (YoY, %)",
      color = NULL, linewidth = NULL, linetype = NULL
    ) +
    theme(panel.grid.minor = element_blank()) +
    facet_wrap(~ type)

eval_cumse_plot_comb <- forecast_metrics %>%
  mutate(class = if_else(family == "benchmark", glue("Benchmark ({model})"), type)) %>%
  select(model, type, class, frcst) %>%
  unnest(frcst) %>%
  mutate(cumse = cumsum(error^2), .by = model) %>%
  ggplot(aes(x = date)) +
    geom_line(aes(y = cumse, group = model, color = class, linewidth = class, linetype = class)) +
    scale_color_manual(
      values = c(
        "Benchmark (AR2)" = "#4c78a8",
        "Benchmark (VAR2)" = "#f58518",
        "Benchmark (VAR2d)" = "#f58518",
        "ARDL" = "#72b7b2",
        "ARDL_ARMA" = "#e45756",
        "ARMAX" = "violet"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "Benchmark (AR2)" = 1.5,
        "Benchmark (VAR2)" = 1.5,
        "Benchmark (VAR2d)" = 1.5,
        "ARDL" = 0.7,
        "ARDL_ARMA" = 0.7,
        "ARMAX" = 0.7
      )
    ) +
    scale_linetype_manual(
      values = c(
        "Benchmark (AR2)" = "solid",
        "Benchmark (VAR2)" = "solid",
        "Benchmark (VAR2d)" = "dashed",
        "ARDL" = "solid",
        "ARDL_ARMA" = "solid",
        "ARMAX" = "solid"
      )
    ) +
    labs(
      title = "Cumulative squared forecast errors",
      x = NULL,
      y = "Cumulative squared error",
      color = NULL, linewidth = NULL, linetype = NULL
    ) +
    theme(panel.grid.minor = element_blank())

# eval_cumse_plot_comb + facet_wrap(~ type)

eval_cumse_exc_plot_comb <- forecast_metrics %>%
  mutate(class = if_else(family == "benchmark", glue("Benchmark ({model})"), type)) %>%
  select(model, type, class, frcst) %>%
  unnest(frcst) %>%
  filter(date < "2020-01-01" | date > "2022-10-01") %>%
  mutate(cumse = cumsum(error^2), .by = model) %>%
  ggplot(aes(x = date)) +
    geom_line(aes(y = cumse, group = model, color = class, linewidth = class, linetype = class)) +
    scale_color_manual(
      values = c(
        "Benchmark (AR2)" = "#4c78a8",
        "Benchmark (VAR2)" = "#f58518",
        "Benchmark (VAR2d)" = "#f58518",
        "ARDL" = "#72b7b2",
        "ARDL_ARMA" = "#e45756",
        "ARMAX" = "violet"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "Benchmark (AR2)" = 1.5,
        "Benchmark (VAR2)" = 1.5,
        "Benchmark (VAR2d)" = 1.5,
        "ARDL" = 0.7,
        "ARDL_ARMA" = 0.7,
        "ARMAX" = 0.7
      )
    ) +
    scale_linetype_manual(
      values = c(
        "Benchmark (AR2)" = "solid",
        "Benchmark (VAR2)" = "solid",
        "Benchmark (VAR2d)" = "dashed",
        "ARDL" = "solid",
        "ARDL_ARMA" = "solid",
        "ARMAX" = "solid"
      )
    ) +
    labs(
      title = "Cumulative squared forecast errors (excluding COVID)",
      x = NULL,
      y = "Cumulative squared error",
      color = NULL, linewidth = NULL, linetype = NULL
    ) +
    theme(panel.grid.minor = element_blank())

# eval_cumse_exc_plot_comb + facet_wrap(~ type)

eval_act_frcst_plot <- grid.arrange(grobs = discard(forecast_metrics$act_frcst_plot, is.null))
ggsave("output/figures/eval_act_frcst.png", eval_act_frcst_plot, width = 30, height = 15, dpi = 300)

eval_cumse_plot <- grid.arrange(grobs = discard(forecast_metrics$cumse_plot, is.null))
ggsave("output/figures/eval_cumse.png", eval_cumse_plot, width = 30, height = 15, dpi = 300)

eval_cumse_exc_plot <- grid.arrange(grobs = discard(forecast_metrics$cumse_exc_plot, is.null))
ggsave("output/figures/eval_cumse_exc.png", eval_cumse_exc_plot, width = 30, height = 15, dpi = 300)
```

```{r forecast_eval_plots_comb, fig.width = 16, fig.height = 12}
eval_act_frcst_plot_comb
# ggsave("output/figures/eval_act_frcst_comb.png", eval_act_frcst_plot_comb, width = 16, height = 12, dpi = 300)
{eval_act_frcst_plot_comb + labs(title = NULL)} %>%
ggsave("output/figures/eval_act_frcst_comb.png", ., width = 8.5, height = 4.6, dpi = 300)

eval_cumse_plot_comb
# ggsave("output/figures/eval_cumse_comb.png", eval_cumse_plot_comb, width = 16, height = 12, dpi = 300)
{eval_cumse_plot_comb + labs(title = NULL)} %>%
ggsave("output/figures/eval_cumse_comb.png", ., width = 6, height = 3.5, dpi = 300)

eval_cumse_exc_plot_comb
# ggsave("output/figures/eval_cumse_exc_comb.png", eval_cumse_exc_plot_comb, width = 16, height = 12, dpi = 300)
{eval_cumse_exc_plot_comb + labs(title = NULL)} %>%
ggsave("output/figures/eval_cumse_exc_comb.png", ., width = 6, height = 3.5, dpi = 300)
```

## Comment

The recursive 1-step-ahead forecasts and cumulative squared errors show that the
main difficulty is COVID: most models miss the 2020Q2 contraction and then under-
or over-adjust during the 2021 rebound.
Before COVID, forecast errors are relatively small across models. During COVID,
the benchmark AR(2), VAR(2), and plain ARDL models accumulate large errors, while
errors for the ARDL-ARMA and ARMAX specifications rise more slowly, as they better
track the rebound and normalization. This suggests that modelling residual dynamics
is important for this forecasting exercise.
We also perform a "normal times" comparison, excluding the 2020Q1-2022Q4 COVID period,
which also supports ARDL-ARMA dominance, so the forecast gains are not only driven
by extreme pandemic observations.

The MSFE and RMSFE rankings favour the ARDL-ARMA and ARMAX models over the AR(2)
and VAR(2) benchmarks. The best full-sample performer is ARDL_ARMA5010, improving
on the AR(2) by 44.3% in MSFE terms. A more economically complete alternative is
ARDL_ARMA5250 because it includes GDP growth, inflation, the change in the cash
rate, and unemployment without significantly sacrificing forecasting performance.

The best ARDL model is ARDL5201, a simple single-lag model including inflation,
only marginally improving on an AR2 model. And for the ARMAX models, the best balance
between full-sample and "normal-times" forecasting is given by ARMAX1450.

### Forecast tests

The preferred ARDL-ARMA and ARMAX models generally pass the forecast bias and forecast-error
autocorrelation tests. This means their forecast errors are not clearly different
from zero on average and do not show strong remaining serial correlation. By contrast,
some benchmark and plain ARDL models show weaker forecast-error diagnostics.

Although MSFE reductions relative to the AR2 are large, the Diebold-Mariano and
HAC loss-difference tests are only significant at the 10% level for the best models.
Therefore, the statistical evidence for forecasting improvements is marginal. This
is not surprising because the evaluation sample only has 40 quarters, and a large
share of the loss difference comes from a few extreme COVID observations.

# Finalists

```{r finalists, fig.width = 16, fig.height = 12}
# Finalist evaluation
# Show everything

finalists <- c("AR2", "VAR2", "ARDL_ARMA5010", "ARDL_ARMA5250", "ARDL5201", "ARMAX1450")

forecast_finalists <- model_diagnostics %>%
  filter(model %in% finalists) %>%
  left_join(select(forecast_metrics, -any_of(names(model_diagnostics)), model), by = "model")

forecast_finalists %>%
  select(-c(type, regressors, controls, fit, frcst, ends_with("plot"))) %>%
  kable(
    digits = 3,
    caption = "Finalist performance"
  )

fin_act_frcst_plot_comb <- forecast_finalists %>%
  select(model, type, frcst) %>%
  unnest(frcst) %>%
  ggplot(aes(x = date)) +
    geom_line(aes(y = actual, color = "Actual", linewidth = "Actual", linetype = "Actual")) +
    geom_line(aes(y = forecast, color = model, linewidth = model, linetype = model)) +
    scale_color_manual(
      values = c(
        "Actual" = "#111111",
        "AR2" = "#4c78a8",
        "VAR2" = "#f58518",
        "ARDL5201" = "#72b7b2",
        "ARDL_ARMA5010" = "#e45756",
        "ARDL_ARMA5250" = "#e45756",
        "ARMAX1450" = "violet"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "Actual" = 1.5,
        "AR2" = 1.5,
        "VAR2" = 1.5,
        "ARDL5201" = 1,
        "ARDL_ARMA5010" = 1,
        "ARDL_ARMA5250" = 1,
        "ARMAX1450" = 1
      )
    ) +
    scale_linetype_manual(
      values = c(
        "Actual" = "solid",
        "AR2" = "solid",
        "VAR2" = "solid",
        "ARDL5201" = "solid",
        "ARDL_ARMA5010" = "solid",
        "ARDL_ARMA5250" = "dashed",
        "ARMAX1450" = "solid"
      )
    ) +
    labs(
      title = "Recursive 1-ahead forecasts",
      x = NULL,
      y = "GDP growth (YoY, %)",
      color = NULL, linewidth = NULL, linetype = NULL
    ) +
    theme(panel.grid.minor = element_blank())

fin_cumse_plot_comb <- forecast_finalists %>%
  select(model, type, frcst) %>%
  unnest(frcst) %>%
  mutate(cumse = cumsum(error^2), .by = model) %>%
  ggplot(aes(x = date)) +
    geom_line(aes(y = cumse, color = model, linewidth = model, linetype = model)) +
    scale_color_manual(
      values = c(
        "AR2" = "#4c78a8",
        "VAR2" = "#f58518",
        "ARDL5201" = "#72b7b2",
        "ARDL_ARMA5010" = "#e45756",
        "ARDL_ARMA5250" = "#e45756",
        "ARMAX1450" = "violet"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "AR2" = 1.5,
        "VAR2" = 1.5,
        "ARDL5201" = 1,
        "ARDL_ARMA5010" = 1,
        "ARDL_ARMA5250" = 1,
        "ARMAX1450" = 1
      )
    ) +
    scale_linetype_manual(
      values = c(
        "AR2" = "solid",
        "VAR2" = "solid",
        "ARDL5201" = "solid",
        "ARDL_ARMA5010" = "solid",
        "ARDL_ARMA5250" = "dashed",
        "ARMAX1450" = "solid"
      )
    ) +
    labs(
      title = "Cumulative squared forecast errors",
      x = NULL,
      y = "Cumulative squared error",
      color = NULL, linewidth = NULL, linetype = NULL
    ) +
    theme(panel.grid.minor = element_blank())

fin_cumse_exc_plot_comb <- forecast_finalists %>%
  select(model, type, frcst) %>%
  unnest(frcst) %>%
  filter(date < "2020-01-01" | date > "2022-10-01") %>%
  mutate(cumse = cumsum(error^2), .by = model) %>%
  ggplot(aes(x = date)) +
    geom_line(aes(y = cumse, color = model, linewidth = model, linetype = model)) +
    scale_color_manual(
      values = c(
        "AR2" = "#4c78a8",
        "VAR2" = "#f58518",
        "ARDL5201" = "#72b7b2",
        "ARDL_ARMA5010" = "#e45756",
        "ARDL_ARMA5250" = "#e45756",
        "ARMAX1450" = "violet"
      )
    ) +
    scale_linewidth_manual(
      values = c(
        "AR2" = 1.5,
        "VAR2" = 1.5,
        "ARDL5201" = 1,
        "ARDL_ARMA5010" = 1,
        "ARDL_ARMA5250" = 1,
        "ARMAX1450" = 1
      )
    ) +
    scale_linetype_manual(
      values = c(
        "AR2" = "solid",
        "VAR2" = "solid",
        "ARDL5201" = "solid",
        "ARDL_ARMA5010" = "solid",
        "ARDL_ARMA5250" = "dashed",
        "ARMAX1450" = "solid"
      )
    ) +
    labs(
      title = "Cumulative squared forecast errors (excluding COVID)",
      x = NULL,
      y = "Cumulative squared error",
      color = NULL, linewidth = NULL, linetype = NULL
    ) +
    theme(panel.grid.minor = element_blank())
```

```{r finalist_plot_comv, fig.width = 16, fig.height = 12}
fin_act_frcst_plot_comb
fin_cumse_plot_comb
fin_cumse_exc_plot_comb
```
